---
title: "Marketing Spend on Customer Purchase Behaviour"
author: "Siyuan Yin 32876793"
format:
  html: 
     number-sections: true
     toc: true
  pdf: 
     number-sections: true
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false
# load packages 
library(tidyverse)
library(knitr)
library(here)

#Upload the data by relative path
marketing_raw_data<-read.csv(here("data/32876793_SiyuanYin.csv"))
```

# Introduction {#sec-Introduction}
Marketing campaigns are a key promotional tool for driving sales. This report examines the relationship between the costs of personalized marketing campaigns and customer purchase amount over the last quarter. The aim is to evaluate the effectiveness of these marketing activities and provide evidence to inform future decision-making.

# Research Question {#sec-Research_Question}
This study investigates whether higher **marketing_spend** is associated with increased **purchase_amount** at the customer level. The second objective is to examine which segment has the highest conversion efficiency from marketing to purchase.

# Dataset Introduction {#sec-Data_Introduction}

**Describing data**.  
  The dataset records customer information and measures how customers are influenced by marketing campaigns. It includes details such as a unique identifier, sign-up date, customer segment, marketing expenditure, purchase behaviour, visit frequency, and additional information. Overall, this dataset provides a comprehensive record of customer activity and engagement, which can be used to analyse the relationship between marketing campaigns and customer purchasing behaviour.

**Data Clean**.  
The data were cleaned by following steps. The `signup_date` variable was converted into a POSIXct format. Observations with missing values in variables (`purchase_amount` and `marketing_spend`) were removed. Missing values in the `notes` column were replaced by NA. 
```{r}
#| label: data_clean
# save the clean data to marketing_data
marketing_data<- marketing_raw_data|>
#delete rows that "purchase_amount" column are missing
  drop_na(purchase_amount) |>
    drop_na(marketing_spend) |>
# convert "signup_date" to POSIXct structure.(date) (y-year,m-month,d-day)
# can ignore the warning message, cause it will not recognize "N/A" in the "signup_date", but will transfer "N/A" to NA
  mutate(signup_date=
  parse_date_time(
    signup_date,orders=c("ymd","dmy","mdy")),
#replace missing value by NA in "notes" column 
notes=na_if(notes,""),
  notes=na_if(notes,"N/A"))
```


## Dataset Description {#sec-Dataset_Description}

The dataset contains **`r nrow(marketing_data)`** observations, and **`r ncol(marketing_data)`** variables. The dataset includes identifier, signup date, marketing cost, purchasing behavior, visit, some additional information and categories of customer that together describe the business campaigns. The **customer_id** in string type identifies each individual customer, while **signup_date** in date type records the date of customer sign-up. The numerical variables, which include **marketing_spend**, **visits**, and **purchase_amount** record the marketing cost, visit times, and customer purchase amount, respectively. The **segment** variable in string type record each category of customers into, such as Consumer, SMB, and Enterprise. Finally, the **notes** variable in string type provides additional information.   


**@fig-segment_count** illustrates the distribution of customers across the three segments. The x-axis represents the customer segments (Consumer, Enterprise, and SMB), while the y-axis shows the total number of customers in each group. As shown in @fig-segment_count, the consumer segment has the largest number of customers, followed by SMB, while Enterprise represents the smallest group.
```{r}
#| label: fig-segment_count
#| fig-cap: "number of customers by segment"
#| fig-width: 5
#| fig-height: 3

marketing_data |>
ggplot(aes(x = segment)) +
#each segment has different color
  geom_bar(aes(fill=segment)) +
  labs(
    x = "Segment",
    y = "Count",
  )
```


**@tbl-segment_summary** below is a summary of categorical data organized by segment. The table shows the total number of customers in each category, as well as the average marketing spend, average purchase amount, and average number of visits. @tbl-segment_summary shows that marketing spend is highest for SMB, followed by Consumer, and lowest for Enterprise. Purchase amount is highest for SMB, followed by Consumer, and lowest for Enterprise. The order of visits is the same as above. This indicates that SMB customers are generally more engaged and generate higher spending.
```{r}
#| label: tbl-segment_summary
#| tbl-cap: "Summary for each segment"

  marketing_data |>
#summary by each segment   
    group_by(segment) |>
    summarise(
   Count = n(),
#round(x,2) means only save 2 decimal places
   "Avg_marketing"=round(mean(marketing_spend),2),
   "Avg_purchase"=round(mean(purchase_amount),2),
   "Avg_visits"=round(mean(visits),2)
     ) |>
    kable()
```


# Results {#sec-Results}

```{r}
#| label: check the linear model
#| include: false

#The regression results are significant (p-value < 0.001), there is a significant positive relationship between marketing spend(x) and purchase amount(y).
lm_model <- lm(purchase_amount ~ marketing_spend, data = marketing_data)
summary(lm_model)
```

**@fig-linear_regression** shows a **positive linear association** between marketing_spend and purchase_amount. This indicates that higher marketing expenditure is generally associated with higher customer spending. However, the dispersion of data points around the regression line suggests that there is substantial individual variation, implying that marketing spend is not the only factor influencing purchasing behaviour.
```{r}
#| label: fig-linear_regression
#| fig-cap: "Relationship between marketing_spend and purchase_amount"
#| fig-width: 5
#| fig-height: 3

#plot the regression
ggplot(marketing_data, aes(x = marketing_spend, y = purchase_amount)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)+
  labs(
    x = "Marketing spend $",
    y = "purchase amount $",
  )
```


**@fig-boxplot** displays the marketing ratio for each of these three customer segments(ratio=divide purchase amount by marketing cost). Most data points fall between **0.5** and **1.5**, and there are no significant differences among the three segments. In terms of the median, the Enterprise segment is higher than the other two segments, indicating that the **Enterprise segment** typically has a slightly more efficient marketing conversion ratio. Additionally, an examination of the outliers reveals that the Enterprise segment has some high outliers, suggesting that in certain cases, this category may achieve exceptionally high returns on marketing investment. In summary, the **Enterprise segment** demonstrates higher marketing efficiency and purchasing potential, but also exhibits relatively significant volatility.
```{r}
#| label: fig-boxplot
#| fig-cap: "The conversion efficiency of marketing expenditures"
#| fig-width: 5
#| fig-height: 4

 #creat new dataframe that contain ratio
 marketing_ratio<- marketing_data |>
# ratio means every 1 marketing spend increase, how much purchase increase
  mutate(ratio=round((purchase_amount/marketing_spend),2)) |>
  select(segment,ratio)

#plot the marketing_ratio by each segment
marketing_ratio |>
  ggplot(aes(x = segment, y = ratio, fill = segment)) +
  geom_boxplot() +
 #control the scale of y-axis (0-14)
  coord_cartesian(ylim = c(0, 14))+
  labs(
    x = "Customer Segment",
    y = "Marketing Efficiency ratio"
  )
```


Overall, returning to @sec-Research_Question, the results indicate a positive linear association between marketing spend and purchase amount. This suggests that higher marketing spend is generally associated with increased purchase amount. In addition, the Enterprise segment demonstrates slightly higher conversion efficiency from marketing investment to purchase, although with greater variability.

